---
title: "Practical Part Appplied Biostatistics II"
author: "Stephan Peischl"
date: "3/15/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Exercise 7

A researcher collected daffodils from four sides of a building and from an open area nearby.
She wondered whether the average stem length of a daffodil depends on its location. The data set is
available as daffodils.csv from ILIAS.
a) State the null hypothesis of an ANOVA model for this problem in words and as a formula.

The null hypothesis is that all locations are the same and there is no differnce in growth.

Formula:
Yij = observation j in group i
ai= groupeffect of the locations, East = intercept

$Y_{ij} = \mu + a_i + E_{ij}$   with  $E_{ij} ~ N(\mu,\sigma)$

H0: aNorth = aOpen = aWest = aSouth = 0


b) Based on the boxplot, does it appear that the null hypothesis is true?

The a for the location Open seems to be different from the others


c) Fit an ANOVA model to the data and test the null hypothesis from a) on a significance level of 10%.

```{r}
flower_data <-read.table('daffodils.csv', header = T, sep = ',')

flower_anova <-lm(Length ~ Side, data = flower_data)

summary(flower_anova)
```

The side Open is different from the others with a p-value of 0.00811.


d) Does the ANOVA model fit well to the data? Perform a residual analysis.

```{r}
plot(fitted(flower_anova), flower_anova$residuals)
qqnorm(flower_anova$residuals)

library(car)
qqPlot(flower_anova$residuals)
```

The residuals seem do not show any major skew. The model seems to fit well. 


e) Which locations (sides of the building and open area) are not significantly different on a 5% level?
Use Bonferroni adjusted pairwise t-tests.

```{r}
pairwise.t.test(flower_data$Length, flower_data$Side, p.adjust.method = 'bonferroni')
help("pairwise.t.test")
```
The t-test reveals that all locations execept Open and South don't differ from each other
given a significance level of 5%.

# Exercise 8

A researcher studied the flexibility of women after taking different sports courses. The

exibility was measured by the spinal extension, a measure of how far the women could bend her
back.
The ANOVA table of the data set looks as follows:

Analysis of Variance Table

Response: SpineExtension

Df Sum Sq Mean Sq F value Pr(>F)

Activity 2 7.0357 3.5178 6.0667 0.006882 **

Residuals 26 15.0764 0.5799

\-\-\-

Signif. codes:

0 \*\*\* 0.001 \*\* 0.01 \* 0.05 . 0.1  \` \`  1

a) How many groups (sports activities) were in the experiment? How many women participated?

Activity has 2 degrees of freedom. We use g-1 for the ANOVA so the original amount of
groups=g was 3.
The residuals have 26 degrees of freedom. We use n-g for the ANOVa so the original 
n was 26+3=29


b) What can you say on a 5% level about the null hypothesis that all sport courses lead to the same flexibility?

We see that the p-value is clearly below the significance level, so the sports are not all 
the same. We can not say which ones are different

c) What is the pooled standard deviation, spool?

spool^2 = the mean square of errors
-> spool = sqrt(mean square of errors) = sqrt(0.5799)
